Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method

### 2023-09-01 manu-draft.qmd
added tassone ref

A unified model to explain the large variability of metabolic rates within and across estuaries is clearly lacking as a critical tool for understanding metabolic processes. Bayesian techniques have not been fully explored in estuarine applications despite their potential to address the above [although see @ciavatta2008]. Bayesian techniques allow for incorporation of prior knowledge about model parameters and robust estimation of parameter probability distributions [@Holtgrieve10] using highly efficient computational algorithms [@Grace15; @winslow2016]. These techniques are very powerful because they not only estimate P and R as standard open-water methods do, but they retrieve estimates and their uncertainties of parameters related to gas exchange, P, and R (e.g., the initial slope of the photosynthesis--irradiance curve), which reveals quantitative information about the mechanisms of metabolism and gas exchange. A potentially useful application developed for streams is the BAyesian Single-station Estimation (BASE) method [@Grace15]. BASE has been demonstrated to accurately determine metabolic parameters and their uncertainty using continuous monitoring data that are similar to those available at many estuarine locations. As such, modification of BASE to include model characterizations more appropriate for estuaries, while maintaining the fundamental Bayesian approach, could prove useful.

>>>>

A unified model to explain the large variability of metabolic rates within and across estuaries is clearly lacking as a critical tool for understanding metabolic processes. Bayesian techniques have only been explored in limited estuarine applications despite their potential to address the above \[ @ciavatta2008; @tassone2019\]. Bayesian techniques allow for incorporation of prior knowledge about model parameters and robust estimation of parameter probability distributions [@Holtgrieve10] using highly efficient computational algorithms [@Grace15; @winslow2016]. These techniques are very powerful because they not only estimate P and R as standard open-water methods do, but they retrieve estimates and their uncertainties of parameters related to gas exchange, P, and R (e.g., the initial slope of the photosynthesis--irradiance curve), which reveals quantitative information about the mechanisms of metabolism and gas exchange. A potentially useful application developed for streams is the BAyesian Single-station Estimation (BASE) method [@Grace15]. BASE has been demonstrated to accurately determine metabolic parameters and their uncertainty using continuous monitoring data that are similar to those available at many estuarine locations. As such, modification of BASE to include model characterizations more appropriate for estuaries, while maintaining the fundamental Bayesian approach, could prove useful.
